---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nisrarr

<!-- badges: start -->
[![R-CMD-check](https://github.com/MarkPaulin/nisrarr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MarkPaulin/nisrarr/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/MarkPaulin/nisrarr/graph/badge.svg)](https://app.codecov.io/gh/MarkPaulin/nisrarr)
[![CRAN status](https://www.r-pkg.org/badges/version/nisrarr)](https://CRAN.R-project.org/package=nisrarr)
<!-- badges: end -->

`nisrarr` is a package for accessing data from the 
[NISRA data portal](https://data.nisra.gov.uk) directly from R.

## Installation

Install nisrarr from CRAN with:

``` r
install.packages("nisrarr")
```

Alternatively, you can install the development version of nisrarr from 
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("MarkPaulin/nisrarr")
```

## Example

`nisra_search()` lets you search for a dataset using keywords or variable names,
and shows information like the last time the dataset was updated:

```{r search}
library(nisrarr)
head(nisra_search(keyword = "claimant"))
```

`nisra_read_dataset()` can be used to download a dataset from the NISRA data 
portal as a data-frame:

```{r data}
claimant_count <- nisra_read_dataset("CCMLGD")
head(claimant_count)
```

The data portal also provides various types of metadata, which can be accessed 
using `get_metadata()` or `get_metadata_field()`:

```{r metadata}
get_metadata(claimant_count)

get_metadata_field(claimant_count, "contact")
```

You can also access boundaries from the data portal. This isn't available for 
all datasets, so your mileage may vary:

```{r bounds}
library(dplyr)
library(ggplot2)

claimant_count |> 
  nisra_get_boundaries() |> 
  filter(Statistic == "Percentage Claimant Count", Month == max(Month)) |> 
  ggplot(aes(fill = value)) +
  geom_sf() +
  scale_fill_binned(
    type = "viridis", 
    n.breaks = 6, 
    name = "Percentage Claimant Count",
    labels = scales::label_percent(scale = 1)
  )
```


## Help wanted / things that need done

There isn't much error handling at the moment. If you're using this package and 
get an error message, please open an issue!

## Inspiration

This package is heavily based on [csodata](https://github.com/CSOIreland/csodata), which wraps an identical API. I also learnt a lot from reading the code for [nomisr](https://github.com/ropensci/nomisr).
